---
title: "Arguments"
author: "Christopher Dory"
date: "2025-11-18"
output: html_document
---
```{r setup, include=FALSE,  message = FALSE, error = FALSE, warning = FALSE}
## IMPORTANT 
## POINT THIS TO YOUR DIRECTORY
knitr::opts_knit$set(echo = TRUE,
                      root.dir = getwd())
```

```{r setup2, include=FALSE, message = FALSE, error = FALSE, warning = FALSE}
#===========================================================================================
# Requirements Load
#===========================================================================================
library(sf)
library(sp)
library(raster)
library(colorspace)
library(terra)
library(lubridate)
library(stringr)
library(here)
library(DT)
options(scipen = 100)
```

# Variable documentation

<b><font size = "3">streams</font></b>: *sf object*, series of linestrings describing the streams in the domain.
<br/>
<br/>
<b><font size = "3">streams_are_points</font></b>: *boolean* default *FALSE*, describes whether the streams being passed are linestrings, in which the program will transform them to points for calculations, or whether they are pre-processed by the user to points. An argument of *FALSE* equates to telling the program you are passing linestrings.
<br/>
<br/>
<b><font size = "3">stream_id_key</font></b>: *char* default *NULL*, if streams are already points this argument describes an ID column in that point set. The ID describes what reach each point belongs to. IDs must be sequentially numeric (i.e. 1,2,3,4...).
<br/>
<br/>
<b><font size = "3">wells</font></b>: *sf object*, series of points that represent the wells in the domain.
<br/>
<br/>
<b><font size = "3">wells_id_key</font></b>: *char* default *NULL*, describes a column in that point set where an ID can be found in the wells data. The ID describes what reach each point belongs to. IDs must be sequentially numeric (i.e. 1,2,3,4...).
<br/>
<br/>
<b><font size = "3">pumping</font></b>: *matrix*, a matrix where the number of rows equals the number of wells, and the number of columns equals the number of timesteps. Each row, column pair contains a pumping rate for that timestep. For example, if the timesteps are in days and m[1,1] = 100 this is equivalent to a pumping of 100 units on day 1 for well 1.
<br/>
<br/>
<b><font size = "3">subwatersheds</font></b>: *sf object* default *NULL*, if the proximity criteria is set to 'adjacent' or 'adjacent+expanding' this argument is necessary. It describes what subwatersheds are to be used in the proximity criteria assignment. For each well streams within the subwatershed of the well or adjacent subwatersheds are considered to be effected by the wells pumping.
<br/>
<br/>
<b><font size = "3">influence_radius</font></b>: *numeric* default *NULL*, if the proximity criteria is set to 'local area', 'expanding', or 'adjacent+expanding' this argument is necessary. It describes the radius around each well that they influence. Streams within this radius are considered to be effected by the wells pumping.
<br/>
<br/>
<b><font size = "3">proximity_criteria</font></b>: *char* default *'local area'*, describes what proximity criteria to use when assigning what streams are effected by the pumping of a given well. Can take arguments of *'local area', 'whole domain', 'adjacent', 'adjacent+expanding', 'expanding'*. In this program *expanding* and *local area* are equivalent, however the number passed to *influence_radius* by the user should necessarily be different given what they represent.
<br/>
<br/>
<b><font size = "3">apportionment_criteria</font></b>: *char* default *'inverse distance'*, describes how to assign for each well what fraction of its pumping goes to each stream it effects. The sum of the fractions, no matter the argument passed here, will always total to 1. Accepts *'inverse distance', 'inverse distance squared', 'web', 'web squared', 'thiessen polygon'*.
<br/>
<br/>
<b><font size = "3">analytical_model</font></b>: *char* default *'glover'*, describes what analytical model to use for the calculation of stream depletions. Accepts *'glover', 'hunt', 'hantush'*. *'Hunt'* and *'hantush'* models are equivalent under the special case that $\lambda = 2\frac{T}{L}$ according to Reeves (2008).
<br/>
<br/>
<b><font size = "3">stream_depletion_output</font></b>: *char* default *'volumetric'*, describes whether the output of the program should be A) the volume of depletions experienced by each reach or B) depletions as a fraction of the pumping contributing to those depletions. Accepts *'volumetric' or 'fractional'*
<br/>
<br/>
<b><font size = "3">lagged_depletions_start_time</font></b>: *numeric* default *0.01*, if *stream_depletion_output* is set to *'fractional'* this describes where to start summing pumping for the denominator of depletions/pumping. As pumping does not immediately effect stream depletions, this is saying 'find the average time at which 1% of depletions occur for pumping on this reach'. Pumping is then summed between [:t-lagged_depletions_start_time]. Accepts any value between 0 and 1, however it should not be larger than *lagged_depletions_end_time*.Average time is calculated as a weighted average, where wells are weighted by proximity. This prevents a sudden increase in proximate well pumping to not enter the denominator until an inappropriate time, blowing up the fraction to infinity.
<br/>
<br/>
<b><font size = "3">lagged_depletions_end_time</font></b>: *numeric* default *0.99*, if *stream_depletion_output* is set to *'fractional'* this describes where to sum pumping for the denominator of depletions/pumping. As pumping does not effect stream depletions indefinitely, this is saying 'find the average time at which 99% of depletions occur for pumping on this reach'. Pumping is then summed between [t-lagged_depletions_end_time:]. Accepts any value between 0 and 1, however it should not be smaller than *lagged_depletions_start_time*. Average time is calculated as a weighted average, where wells are weighted by proximity. This prevents a sudden increase in proximate well pumping to not enter the denominator until an inappropriate time, blowing up the fraction to infinity.
<br/>
<br/>
<b><font size = "3">data_out_dir</font></b>: *char* default *getwd()*, describes where to place the output data.
<br/>
<br/>
<b><font size = "3">diag_out_dir</font></b>: *char* default *getwd()*, describes where to place the diagnostic data (currently limited to log.txt).
<br/>
<br/>
<b><font size = "3">suppress_loading_bar</font></b>: *boolean* default *TRUE*, describes whether to suppress loading bars in console. Value of *TRUE* does not give loading bars.
<br/>
<br/>
<b><font size = "3">suppress_console_messages</font></b>: *boolean* default *TRUE*, describes whether to supress user messages to console. Value of *TRUE* does not give console messages.
<br/>
<br/>
<b><font size = "3">stor_coef_key</font></b>: *char* default *'Stor'*, describes the name of the column in the well data that stores the storage coefficient data.
<br/>
<br/>
<b><font size = "3">transmissivity_key</font></b>: *char* default *'Tr'*, describes the name of the column in the well data that stores transmissivity data ($Kb$).
<br/>
<br/>
<b><font size = "3">leakance_key</font></b>: *char* default *NULL*, describes the name of the column in the stream data that stores leakance data, where leakance is a product of the conductivity of the aquifer multiplied by the streambed clogging layer thickness divided by the streambed clogging layer conductivity $\frac{Kb^\text{'}}{K^\text{'}}$. Argument is required if *analytical_model* is set to *'hantush'*. Definition taken from Reeves (2008) referencing Hantush (1965).
<br/>
<br/>
<b><font size = "3">lambda_key</font></b>: *char* default *NULL*, describes the name of the column in the stream data that stores lambda data, where lambda is a product of the width of the river multiplied by the conductivity of the streambed clogging layer divided by the thickness of the streambed clogging layer $w_{r}*\frac{K_{r}}{b_{r}}$. Argument is required if *analytical_model* is set to *'hunt'*. Definition taken from Zipper et al. (2019) referencing Hunt (1999).
<br/>
<br/>
<b><font size = "3">prec</font></b>: *numeric* default *80*, describes how many bits of precision to use in the Rmpfr package when calculating exponential values. Required for the *'hunt'* and *'hantush'* analytical models.
Exponential values in the *'hunt'* and *'hantush'* analytical models can often return **Inf** using base numerics, use of *prec* in Rmpfr alleviates this. Higher *prec* values will slow down processing. Default taken from streamDepletr package by Zipper (2019).
<br/>
<br/>
<br/>
<br/>

# References

Glover and Balmer (1954) https://doi.org/10.1029/TR035i003p00468 <br/>
Hantush (1965) https://doi.org/10.1029/JZ070i012p02829 <br/>
Hunt (1999) https://doi.org/10.1111/j.1745-6584.1999.tb00962.x <br/>
Reeves (2008) https://doi.org/10.3133/ofr20081166 <br/>
Zipper (2019) https://cran.r-project.org/web/packages/streamDepletr/index.html <br/>
Zipper et al. (2019) https://doi.org/10.1029/2018WR024403 <br/>





